---
title: "Penguins Assignment Question 4"
date: "2022-12-06"
output: 
 pdf_document: default
 html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Loading the packages
```{r}
library(ggplot2)
library(palmerpenguins)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(dplyr))
library(tidyr)
library(ragg)

```

```{r}
#Setting the working directory 
setwd("~/Biology/Year 3/Year 3 R/Computing Assessments/PenguinProject") 
```


##Question 4: Run a statistical test on the Palmer Penguins dataset and produce a figure to explain it.

```{r}
penguins_raw
#want to save this penguins_raw dataset
write.csv(penguins_raw, paste0("data_raw/penguins_raw.csv"))
#Saved penguins raw to a data raw folder

#Now we are defining the cleaning function and what it will do to penguins_raw 
cleaning <- function(penguins_raw) {
  penguins_raw %>% 
    select(-starts_with("delta")) %>%
    select(-Comments)%>%
    clean_names()}

#This is within the cleaning.r function that is saved separately. It removes NA
#values and selects the variables I am including in my data analysis
remove_empty_mass_and_length <- function(penguins_clean){
    penguins_clean %>%
    filter(!is.na(culmen_length_mm)) %>%
    filter(!is.na(body_mass_g)) %>%
    select(body_mass_g, culmen_length_mm, species)
}

source("functions/cleaning.r")
#This specifies where the cleaning function is saved

#We are then applying this function to the penguins_raw dataset
penguins_clean <- cleaning(penguins_raw)
penguins_now_clean <- remove_empty_mass_and_length(penguins_clean)

#Saving penguins clean dataset 
write.csv(penguins_clean, paste0("data_clean/penguins_clean.csv"))

#Saving penguins now clean dataset
write.csv(penguins_now_clean, paste0("data_clean/penguins_now_clean.csv"))

```

Now the data has been cleaned, I am going to test to see if there is a significant correlation between body mass and culmen length. I will use Pearson's correlation coefficient because the two variables are numerical and continuous 

```{r}
penguins_model <- lm(culmen_length_mm ~ body_mass_g, data=penguins_now_clean) 
#Making a linear regression model for body mass and culmen length 

#Can do Pearson's correlation coefficient if the assumption of bivariate normality is not breached. 
plot(penguins_model, which =2)
plot(penguins_model, which=1)
#The assumptions of normality and homogeneity of variance are well-met for this model

#Performing the Pearson's correlation coefficient
cor.test(penguins_clean$body_mass_g, penguins_clean$culmen_length_mm, method=c("pearson"))
#That has given me a p value of 2.2e-16. Suggests that there is a significant correlation 
#between the two variables. Will now visualise this with a scatter plot


#Defining my functions for plotting a graph for all the penguins, and broken down by penguin species
plot_penguin_graph1 <- function(penguins_now_clean){
  penguins_now_clean %>%
    ggplot(
           aes(x=body_mass_g, y=culmen_length_mm))+geom_point()+
    geom_smooth(method="lm", colour="blue")+labs(x="Body Mass (g)", y="Culmen Length (mm)", 
    title="Correlation Between Body Mass and Culmen Length 1")+
    theme_bw()
}   
    
plot_penguin_graph2 <- function(penguins_now_clean){ 
  penguins_now_clean %>%
    ggplot(
           aes(x=body_mass_g, y=culmen_length_mm, colour=species))+ geom_point()+
    scale_colour_manual(values=c("deeppink", "blue", "orange"))+ geom_smooth(method="lm")+
    labs(x="Body Mass (g)", y="Culmen Length (mm)", 
    title="Correlation Between Body Mass and Culmen Length 2", colour="Penguin Species") 
}


source("functions/plotting.r")
#This is where my function is saved

#Calling the plotting function to make penguin_graph1 and penguin_graph2
penguin_graph1 <- plot_penguin_graph1(penguins_now_clean) 
penguin_graph2 <- plot_penguin_graph2(penguins_now_clean)

#Looking at the graphs
penguin_graph1 
penguin_graph2

```

The output of the Pearson's correlation coefficient gave a P value of 2.2e-16. This is less than 0.05, so we can reject the null hypothesis that there is no significant correlation between body mass and culmen length.

When this result is visualised with a scatterplot and regression lines, we see that there is a positive correlation between body mass and culmen length such that larger penguins have a longer culmen. The second scatterplot breaks down this positive correlation for each species, so we can see that the correlation holds true across all species of penguins.

```{r}
#Defining a function for saving penguin graph 1 and penguin graph 2

save_graph1_png <- function(penguins_now_clean,
                                  filename, size, res, scaling){
  agg_png("figures/penguin_graph1.png", 
                    width   =  25, 
                    height  =  20, 
                    units   =  "cm", 
                    res     =  600, 
                    scaling =  1)
  penguin_graph1 <- plot_penguin_graph1(penguins_now_clean)
  print(penguin_graph1)
  dev.off()
}

save_graph2_png <- function(penguins_now_clean,
                                  filename, size, res, scaling){
  agg_png("figures/penguin_graph2.png", 
                    width   =  25, 
                    height  =  20, 
                    units   =  "cm", 
                    res     =  600, 
                    scaling =  1)
  penguin_graph2 <- plot_penguin_graph2(penguins_now_clean)
  print(penguin_graph2)
  dev.off()
}

source("functions/saving.r") 
#this is where my saving function is stored

#Calling the saving function to save penguin graph 1 and penguin graph 2 as png files
penguin_graph1.png <- save_graph1_png(penguins_now_clean)
penguin_graph2.png <- save_graph2_png(penguins_now_clean)

#my graphs have now been saved in the "figures" folder 

```


